Refinements for open automata

### 2023-06-19 2021-Refinement/paper/refinementarticleV2.tex
coherence notations

In previous works \cite{AMHEEMA:2023,wang:03126313} a bisimulation relation was defined for OA and open parameterized hierarchical LTSs. It has already good properties relatively to bisimulation, but refinement relations were not studied. In this article we go further to  define a theory of refinement to OA. The refinement relation we introduce in the paper  is based on the notion of   simulation,  in a similar way to that defined in classical automata theory \cite{Milner:1989,Bellegarde:2000}.  And with the common behaviour-preserving  property:  all the behaviour of the abstract specification must be followed by its (complex) implementation but additional behaviours may exist. However we also ensure that a whole scenario, made of several steps, of the specification can also be simulated by the refined system, which is slightly richer than the traditional simulation relation and which allows us to obtain a compositionality result.

%This article first introduces OA that were defined in  \cite{henrio:01299562}. %,  but compared to previous works, we and additionally defines the composition of OA.
Our contribution in this paper is the definition a refinement relation for OA that has the following characteristics:
\begin{itemize}
\item Classical simulation characterisation but also an additional criteria ensuring that refinement does not introduce deadlocks when following a trace from the simulated automaton.
\item Good properties relatively to composition: we prove that composition preserves the refinement relation.
\item Ability to take into account both composition and transitivity: this is a challenge because composition changes the set of holes of the OA and refinement takes into account the actions of the holes.
\end{itemize}
 The refinement relation is introduced in two steps. First we define a refinement that relates two automata with the same holes, which allows us to focus on the automaton aspect. Second we introduce a relation that relates two automata with different sets of holes, which allows us to take into account the open nature of OA, and to deal with composition. Properties of the refinement are stated and proven on the second, more general version of the relation, they are thus also valid for the first simpler refinement relation.

This paper is organized as follows. Section~\ref{sec:background} recalls the definition of OA and defines their composition. 
We  then  define a refinement relation for OA, first only considering two automata with the same set of holes in Section~\ref{sec:refinement} and generalize it to automata with a different set of holes in Section~\ref{sec:holes}.  Section~\ref{sec:prop} is dedicated to formalize and prove basic properties of the refinement defined  ; here we prove that refinement is a preorder and has nice composability properties. 
In Section~\ref{sec:sota} we review related works, and Section~\ref{sec:ccl} concludes the paper.

@@

\subsection{Relations between open automata}
Establishing semantic equivalences and refinement relations between different OA requires  to compare their states. For this purpose, we suppose that the variables of the two OA are disjoint (a renaming of variables may have to be applied before comparing OA states).
\begin{definition}[Relation on open automata configurations] Suppose $V_1$ and $V_2$ are disjoint.
A relation on configurations of two OA \(\OAg[1]\) and \(\OAg[2]\) is a function \( \mathcal{R}: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\).
\end{definition}
%\TODO{C'est etrange, la formulation "the variables they refer to", comme il n'y a plus de variables d'etat... +1}
The idea is that two states are related depending on the satisfiability of the expression relying their variables i.e., if the variables of the OA verify a certain formula. 
In other words, to each pair of states is attached a boolean formula that may refer to the  variables of each of the two OA, stating whether the two states are related or not.
Additionally, we say that the relation $ \mathcal{R}$ \emph{relates the initial states} of the automata if: \(\sigma_{01} \uplus \sigma_{02} \vdash  \mathcal{R}\mpar{s_{01}, s_{02}}\).

@@

\section{Refinement Relations for automata with the same holes}\label{sec:refinement}

Similarly to FH-bisimulation~\cite{AMHEEMA:2023} we are interested  in finding refinement relations between configurations of two OA that contain variables and holes. When dealing with open systems it is common to define refinement in terms of a simulation relation.
We rely on a classical notion of simulation inspired from~\cite{AMHEEMA:2023}. The idea is that two configurations related by a relation; if one state can do a transition, then the other can do a transition too. 
Like for bisimulation, a refinement relation characterises when two states are related, and this  characterisation is expressed as a predicate on the variables of the two automata.
Refinement defines conditions on a predicate $R$ such that $R\mpar{s_1, s_2}$ is a predicate (possibly involving variables of the automata) that is true when the state $s_1$ of $A_1$ refines the state $s_2$ of $A_2$.

@@

However here we want to build a refinement relation that  also guarantees that no deadlock is introduced when refining the automaton. This property is quite frequent in refinement relation, and referred to as \emph{lack of new deadlocks}~\cite{Kouchnarenko:2007} or \emph{complete simulation}~\cite{sangiorgi:bisim-coind12}.
The notion of deadlock should however be specialised to our OA. Indeed, the possible transitions are better characterised by the guards of the outgoing transitions from a given state, rather than the simple existence of a transition. We thus define a deadlock reduction criterion based on how the outgoing transitions are guarded.
As such, a refinement does not introduce deadlocks if in the conditions where no transition is possible in the refined automaton, no transition are already possible in the more general one. More formally, for any pair of states $s_1$ and $s_2$  we introduce a criterion of the form: 
\[
\forall \mpar{s_1, s_2} \in S_1 \times S_2, V_1 \uplus V_2  \vdash R\mpar{s_1, s_2} \wedge \lnot( \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} )\implies \lnot( \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2}  )
\]

Which can be rewritten as:
\[
\forall \mpar{s_1, s_2} \in S_1 \times S_2, V_1 \uplus V_2  \vdash R\mpar{s_1, s_2} \implies( \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} ) \lor¢ \lnot( \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2}  )
\]

Both statements being equivalent and one of them may reveal more intuitive than the other in different conditions, we will use them interchangeably.
%Let us first explain the reasons why simulation can introduce allow undesired behaviours to emerge and how we want to limit them. 
%The automaton that simulates a specification features any behaviour that contains the behaviour of the specification, or more precisely, for each state that simulates a state of the specification, the simulating automaton must at least feature the behaviour of the specification.
% In this definition, nothing prevents the implementation from 
%
%adding non-determinism and non-deterministically reaching a state that is unrelated to the specification, this state could even be a deadlock even though we only followed transitions that exist in the specification.
%To avoid the appearance of such deadlocks, we will state that, when both the implementation and the specification do the same transition, they stay related.
%This ensures a form of continuity in the comparison in the sense that the refined system is constrained to follow the behaviour of the simulated system along a given scenario. 
%In terms of traces this amounts to some form of deadlock reduction: for each trace followed by the simulated system if the refined system followed the beginning of the trace, it must be able to keep following the trace. 
%Of course, the classical trace view is a simplified view compared to the guards, states and actions from  the holes that exist in OA, but it gives the right intuition.
%
%Next definition states that if $T_2$ is not in a deadlock position then $T_1$ can do a transition but not necessarily with the same input values. This is very weak but sufficient when composed with the def  of simulation.
%\begin{definition}[Preservation along identical transitions]\label{def:preservation} \\
%Let \(\OAg[1]\) and \(\OAg[2]\) be two OAs.
%A relation \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) between  \(\OAg[1]\) and \(\OAg[2]\)
%  is preserved along identical transitions if  it satisfies the following:
%\begin{multline*}
%\forall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} \mpar{t_1, t_2} = \mpar{\OTx{1}{}{1}{1}, \OTx{2}{}{2}{2}} \in \fOT{s_1} \times \fOT{s_2} \\  V_1 \uplus V_2 \uplus \fvars{t_1} \uplus \fvars{t_2} \vdash \hspace{6cm}\\  R\mpar{s_1, s_2} \wedge 	g_1 \wedge g_2 \wedge \alpha_1 = \alpha_2 \wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} \implies  R\mpar{s'_1, s'_2}\psubst{\psi_1 \uplus \psi_2}\\
%\end{multline*}
%\end{definition}
%We thus define the following property that can be understood as follows. Take a specification automaton and its implementation. Consider two related states for the refinement relation. Suppose an identical transition exist in both automata (with same guards and same label), then either the reached states $s_1'$ and $s_2'$ are related by refinement, or there exists another transition of the specification that is also identical and leads to a state $s_1''$ that is related to $s_2'$ in the refinement relation.
%Finally, we should take into account that 1) instead of equality, we should check if there is a state such that both transitions are feasible (the conjunction of guards is satisfiable), and that  2) instead of finding another transition of the specification we might find a family of transitions that cover all the cases of the transition of the implementation. 
We can now state the definition of simulation between OA that have the same set of holes.

\begin{definition}[Hole-equal simulation]\label{def:HoleEqualSim}
Consider two OA with identical set of holes:  \(\OAg[1]\) and \(\OAg[2]\) where  \(J_1 = J_2\), the relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is a hole-\(\symb{equal}\) simulation from $S_1$ to $S_2$ if the following conditions hold (we could denote it $A_1\leq A_2$): 
\item[(1)] \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\)
\item[(2)] \(\forall \mpar{s_1, s_2} \in S_1 \times S_2,\)\vspace{-8pt}
\noindent\begin{multline*}
 \bigsymb{\forall} {t_1} = \openrule
         {
           \beta_{1 j}^{j\in J}, g_1,\psi_1}
         {s_1 \OTarrow {\alpha_1} s_1'} \in \fOT{s_1} .\,~~
\bigsymb{\bigexists} 
\mpar{t_{2x} \!=\! \openrule
         {
           \beta_{2 x j}^{j\in J}, g_{2 x},\psi_{2 x}}
         {s_2 \OTarrow {\alpha_{2 x}} s_{2 x}'}
}^{x \in X}\in \fOT{s_2}.\,~~\\ 
\everymath{\displaystyle}\begin{array}{l}
V_1 \uplus V_2 \uplus  \fvars{t_1} \uplus \bigcup_x\fvars{t_{2x}}  \vdash \hspace{7cm}
\\
 R\mpar{s_1, s_2} \wedge g_1 \implies
 \operatorname*{\bigsymb{\bigvee}}_{x \in X}
\mpar{\begin{array}{l}
			\alpha_{2x} = \alpha_{1} \wedge \bigwedge_\subbox{j \in J} \beta_{2xj} = \beta_{1j} \nwedge\\[12pt]
			 g_{2x} \wedge R\mpar{ s'_{1},s'_{2x}}\psubst{\psi_{2x} \uplus \psi_{1}}
		\end{array}} 
\end{array} 
\end{multline*}
\item [(3)] Deadlock reduction:
\[
\forall \mpar{s_1, s_2} \in S_1 \times S_2, V_1 \uplus V_2  \vdash R\mpar{s_1, s_2} \implies( \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} ) \lor¢ \lnot( \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2}  )
\]

\end{definition}

 %Note also the second part of the conjunction of condition (2) expresses the deadlock reduction between automata (as introduced in Definition \ref{def:dpwd}).   
The first and second conditions coincide with the natural way to prove inductively that an automaton simulates another by starting with the initial state. The third condition ensures that the refinement relation prevents the introduction of deadlocks.
Our definition captures a simple kind of sub-classing of OA with the same holes. It is more expressive than a strict simulation since it matches a transition with a family of transitions. 
With such a relation we are able to check the refinement between two OA with the same level of abstraction but specified differently, for example, by duplicating states, removing transitions,  reinforcing  guards, modifying variables, etc.
We will show in Section~\ref{sec:prop} that this refinement relation has good properties in terms of transitivity, compositionality, and reflexivity.

@@

\begin{example} \label{example:fifo-system-one}
To illustrate the refinement simulation of OA, we consider a variation on the \texttt{prod-cons} example. Namely,  we suppose that  the two processes  \texttt{P} and  \texttt{Q} communicate through a one-place  buffer. Fig. \ref{Fig:SpecOA} shows the OA modelling this simpler version of the system, that we refer to as \texttt{simprod-cons}. 
\begin{figure}[!t]
 \centering
   \includegraphics[width=.90\textwidth]{Figures/FIFOpen2.pdf}
   \caption{  The \texttt{simprod-cons} OA: the system using one-place  buffer.  \label{Fig:SpecOA}}
\end{figure}
We can easily check that this automaton simulates the one of Fig.~\ref{Fig:RefineOA}.
Indeed, one can see that  $R=\{(r_0,s_0) \mapsto \texttt{l}=\texttt{f}, (r_0,s_1) \mapsto \texttt{f}=\texttt{l}+1\% \texttt{N} \}$ is  a candidate simulation 
relation. It follows that  \texttt{prod-cons} $\leq$ \texttt{simprod-cons}. \qed
\end{example}

@@

The refinement relation defined above is insufficient in the setting of composition which is the main advantage of the OA-based approach. Indeed, it should be possible to refine an automaton by filling its hole, providing a concrete view of a part of the application that was not specified originally. 
More generally, it should be possible to relate automata that do not have the same holes because composition is a crucial part of system specification.
%Thus, we believe composition should also be a form of refinement and call this feature \emph{refinement through composition}.
However, filling holes can result in a system with more or less holes than the original system because the plugged subsystem can contain itself many holes.
Next section  defines  a more powerful refinement relation able to reason on automata with different sets of holes.

@@

\section{A Refinement Relation that Takes Holes into Account}\label{sec:holes}

This section  extends the preceding relation to automata where the set of holes is not the same. Indeed, when composing the set of  holes changes, and for example we want to state whether the automaton after composition is a refinement of the original automaton or not.

@@

Unfortunately, the deadlock reduction property in the definition of refinement is not compositional: the composition operator can itself introduce a deadlock. In other words, when filling the hole of two related automata with a third one, even if there is a deadlock reduction between the two original automata, there might not be a deadlock reduction in the composed ones. The same problem may arise when two related automata are composed in the same hole of a third one.

@@

%\begin{definition}[Non-blocking composition] \label{Def:Non-block}
%Let \(A_i = \OAg[i]\) with \( 0 \leq i \leq n\) be a family of open automata.
%Let \(A = A_0\mdbrk{j_i \mapsto A_i \middle| 1 \leq i \leq n}\) and $A= \OAg$.
%
%The composition \(A_0\mdbrk{j_i \mapsto A_i \middle| 1 \leq i \leq n}\) is non-blocking if   \(A \) has a reachability predicate such that, for each reachable configuration, if there is a possible transition in \(A_0\) then there is a possible transition in \(A\):
%\begin{multline*}
% \forall s = \mpar{s_0, s_1,..,s_n} \in S, V \uplus \biguplus_\subbox{t_0 \in \fOT{s_0}} \fvars{t_0} \vdash \\ \left(\reach{A}\mpar{s} \wedge \bigvee_\subbox{t_0 \in \fOT{s_0}} \fguard{t_0} \implies \bigvee_\subbox{t \in \fOT{s}} \fguard{t} \right)
%\end{multline*}
%\end{definition}
\begin{definition}[Non-blocking composition] \label{Def:Non-block}
Let \(A_1 \defobject \OAg[1]\) and \( A_2 \defobject \OAg[2]\) be two OA.
Let $A= \OAg$ be the OA resulting from the composition \(A = A_1\subst{A_2}{k}\). 
The composition \(A_1\subst{A_2}{k}\) is non-blocking if   \(A \) has a reachability predicate such that, for each reachable configuration, if there is a possible transition in \(A_1\) then there is a possible transition in \(A\):
\begin{multline*}
 \forall s = \mpar{s_1, s_2} \in S, V \uplus \biguplus_\subbox{t \in \fOT{s_1}} \fvars{t} \vdash  \left(\reach{A}\mpar{s} \wedge \bigvee_\subbox{t \in \fOT{s_1}} \fguard{t} \implies \bigvee_\subbox{t \in \fOT{s}} \fguard{t} \right)
\end{multline*}
\end{definition}
Like in the definition of refinement (Definition~\ref{def:HoleEqualSim}) we use guards to ensure that the transition can occur. 
In general, one would not want to only consider non-blocking composition as it may reveal a bit restrictive, but it is the best necessary condition that can be identified for compositionality of refinement. It will be used to prove composition theorems in  Section~\ref{sec:prop}. 
Proving refinement relation in absence of non-blocking composition may also checked specifically for the composed automaton under study.
%It is not sufficient to ensure the existence of equivalent transitions in general, but it will be sufficient in the context of refinement.

\subsection{Automaton refinement}

One major challenge in the extension of refinement to different sets of holes is to maintain a form of transitivity while being able to take into account the actions of some of the holes. A naive definition of refinement would ensure that only the holes that are identical in the two OA are taken into account in the simulation. Unfortunately considering all the common holes does not ensure transitivity of the simulation for the following reason. If $A_1$ simulates $A_2$ and $A_2$ simulates $A_3$, and one hole $j$ appears in $A_3$ and in $A_1$ but not in $A_2$ then we have no guarantee on the way $A_1$ and $A_3$ take the actions of this his hole into account, thus  a refinement between and $A_1$ and $A_3$ would require conditions involving actions of the hole $j$ which cannot be ensured. The way we solve this issue is to remember in the simulation relation which holes have been compared. This makes the relation parameterized by a subset of the set of holes that belong to the two automata that we want to take into account.
This way, in the example above, we would have no guarantee on actions the hole $j$ by transitivity but can state a refinement relation with guarantees on the actions of the other holes.

>>>>

In previous works \cite{AMHEEMA:2023,wang:03126313} a bisimulation relation was defined for OA and open parameterized hierarchical LTSs. It has already good properties relatively to bisimulation, but refinement relations were not studied. In this article we go further to  define a theory of simulation for OA. The simulation relation we introduce in the paper  is based on the notion of   simulation,  in a similar way to that defined in classical automata theory \cite{Milner:1989,Bellegarde:2000}.  And with the common behaviour-preserving  property:  all the behaviour of the abstract specification must be followed by its (complex) implementation but additional behaviours may exist. However we also ensure that a whole scenario, made of several steps, of the specification can also be simulated by the refined system, which is slightly richer than the traditional simulation relation and which allows us to obtain a compositionality result.

%This article first introduces OA that were defined in  \cite{henrio:01299562}. %,  but compared to previous works, we and additionally defines the composition of OA.
Our contribution in this paper is the definition a simulation relation for OA that has the following characteristics:
\begin{itemize}
\item Classical simulation characterisation but also an additional criteria ensuring that simulation does not introduce deadlocks when following a trace from the simulated automaton.
\item Good properties relatively to composition: we prove that composition preserves the simulation relation.
\item Ability to take into account both composition and transitivity: this is a challenge because composition changes the set of holes of the OA and simulation takes into account the actions of the holes.
\end{itemize}
 The simulation relation is introduced in two steps. First we define a simulation that relates two automata with the same holes, which allows us to focus on the automaton aspect. Second we introduce a relation that relates two automata with different sets of holes, which allows us to take into account the open nature of OA, and to deal with composition. Properties of the simulation are stated and proven on the second, more general version of the relation, they are thus also valid for the first simpler simulation relation.

This paper is organized as follows. Section~\ref{sec:background} recalls the definition of OA and defines their composition. 
We  then  define a simulation relation for OA, first only considering two automata with the same set of holes in Section~\ref{sec:refinement} and generalize it to automata with a different set of holes in Section~\ref{sec:holes}.  Section~\ref{sec:prop} is dedicated to formalize and prove basic properties of the simulation defined; here we prove that simulation is a preorder and has nice composability properties. 
In Section~\ref{sec:sota} we review related works, and Section~\ref{sec:ccl} concludes the paper.

@@

\subsection{Relations between open automata}
Establishing semantic equivalences and simulation relations between different OA requires  to compare their states. For this purpose, we suppose that the variables of the two OA are disjoint (a renaming of variables may have to be applied before comparing OA states).
\begin{definition}[Relation on open automata configurations] Suppose $V_1$ and $V_2$ are disjoint.
A relation on configurations of two OA \(\OAg[1]\) and \(\OAg[2]\) is a function \( \mathcal{R}: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\).
\end{definition}
%\TODO{C'est etrange, la formulation "the variables they refer to", comme il n'y a plus de variables d'etat... +1}
The idea is that two states are related depending on the satisfiability of the expression relying their variables i.e., if the variables of the OA verify a certain formula. 
In other words, to each pair of states is attached a boolean formula that may refer to the  variables of each of the two OA, stating whether the two states are related or not.
Additionally, we say that the relation $ \mathcal{R}$ \emph{relates the initial states} of the automata if: \(\sigma_{01} \uplus \sigma_{02} \vdash  \mathcal{R}\mpar{s_{01}, s_{02}}\).

@@

\section{Simulation Relations for automata with the same holes}\label{sec:refinement}

Similarly to FH-bisimulation~\cite{AMHEEMA:2023} we are interested  in finding simulation relations between configurations of two OA that contain variables and holes. When dealing with open systems it is common to define simulation in terms of a simulation relation.
We rely on a classical notion of simulation inspired from~\cite{AMHEEMA:2023}. The idea is to consider two configurations related by a relation; if one state can do a transition, then the other can do a transition too. 
Like for bisimulation, a simulation relation characterises when two states are related, and this  characterisation is expressed as a predicate on the variables of the two automata.
Simulation defines conditions on a predicate $\mathcal{R}$ such that $\mathcal{R}\mpar{s_1, s_2}$ is a predicate (possibly involving variables of the automata) that is true when the state $s_1$ of $A_1$ refines the state $s_2$ of $A_2$.

@@

However here we want to build a simulation relation that  also guarantees that no deadlock is introduced when refining the automaton. This property is quite frequent in simulation relation, and referred to as \emph{lack of new deadlocks}~\cite{Kouchnarenko:2007} or \emph{complete simulation}~\cite{sangiorgi:bisim-coind12}.
The notion of deadlock should however be specialised to our OA. Indeed, the possible transitions are better characterised by the guards of the outgoing transitions from a given state, rather than the simple existence of a transition. We thus define a deadlock reduction criterion based on how the outgoing transitions are guarded.
As such, a simulation does not introduce deadlocks if in the conditions where no transition is possible in the refined automaton, no transition were already possible in the more general one. More formally, for any pair of states $s_1$ and $s_2$  we introduce a criterion of the form: 
\[
\forall \mpar{s_1, s_2} \!\in\! S_1 \!\times\! S_2, V_1 \uplus V_2  \vdash \mathcal{R}\mpar{s_1, s_2} \wedge \lnot( \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} )\implies \lnot( \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2}  )
\]

Which can be rewritten as:
\[
\forall \mpar{s_1, s_2} \!\in\! S_1 \!\times\! S_2, V_1 \uplus V_2  \vdash \mathcal{R}\mpar{s_1, s_2} \implies( \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} ) \lor¢ \lnot( \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2}  )
\]

Both statements being equivalent, as each of them may reveal more intuitive than the other in different situations, we  use them interchangeably.
%Let us first explain the reasons why simulation can introduce allow undesired behaviours to emerge and how we want to limit them. 
%The automaton that simulates a specification features any behaviour that contains the behaviour of the specification, or more precisely, for each state that simulates a state of the specification, the simulating automaton must at least feature the behaviour of the specification.
% In this definition, nothing prevents the implementation from 
%
%adding non-determinism and non-deterministically reaching a state that is unrelated to the specification, this state could even be a deadlock even though we only followed transitions that exist in the specification.
%To avoid the appearance of such deadlocks, we will state that, when both the implementation and the specification do the same transition, they stay related.
%This ensures a form of continuity in the comparison in the sense that the refined system is constrained to follow the behaviour of the simulated system along a given scenario. 
%In terms of traces this amounts to some form of deadlock reduction: for each trace followed by the simulated system if the refined system followed the beginning of the trace, it must be able to keep following the trace. 
%Of course, the classical trace view is a simplified view compared to the guards, states and actions from  the holes that exist in OA, but it gives the right intuition.
%
%Next definition states that if $T_2$ is not in a deadlock position then $T_1$ can do a transition but not necessarily with the same input values. This is very weak but sufficient when composed with the def  of simulation.
%\begin{definition}[Preservation along identical transitions]\label{def:preservation} \\
%Let \(\OAg[1]\) and \(\OAg[2]\) be two OAs.
%A relation \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) between  \(\OAg[1]\) and \(\OAg[2]\)
%  is preserved along identical transitions if  it satisfies the following:
%\begin{multline*}
%\forall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} \mpar{t_1, t_2} = \mpar{\OTx{1}{}{1}{1}, \OTx{2}{}{2}{2}} \in \fOT{s_1} \times \fOT{s_2} \\  V_1 \uplus V_2 \uplus \fvars{t_1} \uplus \fvars{t_2} \vdash \hspace{6cm}\\  R\mpar{s_1, s_2} \wedge 	g_1 \wedge g_2 \wedge \alpha_1 = \alpha_2 \wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} \implies  R\mpar{s'_1, s'_2}\psubst{\psi_1 \uplus \psi_2}\\
%\end{multline*}
%\end{definition}
%We thus define the following property that can be understood as follows. Take a specification automaton and its implementation. Consider two related states for the refinement relation. Suppose an identical transition exist in both automata (with same guards and same label), then either the reached states $s_1'$ and $s_2'$ are related by refinement, or there exists another transition of the specification that is also identical and leads to a state $s_1''$ that is related to $s_2'$ in the refinement relation.
%Finally, we should take into account that 1) instead of equality, we should check if there is a state such that both transitions are feasible (the conjunction of guards is satisfiable), and that  2) instead of finding another transition of the specification we might find a family of transitions that cover all the cases of the transition of the implementation. 
We can now state the definition of simulation between OA that have the same set of holes.

\begin{definition}[Hole-equal simulation]\label{def:HoleEqualSim}
Consider two OA with identical set of holes:  \(A_1=\OAg[1]\) and \(A_2=\OAg[2]\) where  \(J_1 = J_2\), the relation on configurations \(\mathcal{R}: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is a hole-\(\symb{equal}\) simulation from $A_1$ to $A_2$ if the following conditions hold : 
\item[(1)] \(\sigma_{01} \uplus \sigma_{02} \vdash \mathcal{R}\mpar{s_{01}, s_{02}}\)
\item[(2)] \(\forall \mpar{s_1, s_2} \in S_1 \times S_2,\)\vspace{-8pt}
\noindent\begin{multline*}
 \bigsymb{\forall} {t_1} = \openrule
         {
           \beta_{1 j}^{j\in J}, g_1,\psi_1}
         {s_1 \OTarrow {\alpha_1} s_1'} \in \fOT{s_1} .\,~~
\bigsymb{\bigexists} 
\mpar{t_{2x} \!=\! \openrule
         {
           \beta_{2 x j}^{j\in J}, g_{2 x},\psi_{2 x}}
         {s_2 \OTarrow {\alpha_{2 x}} s_{2 x}'}
}^{x \in X}\in \fOT{s_2}.\,~~\\ 
%\everymath{\displaystyle}\begin{array}{l}
V_1 \uplus V_2 \uplus  \fvars{t_1} \uplus \bigcup_x\fvars{t_{2x}}  \vdash \hspace{7cm}
\\
 \mathcal{R}\mpar{s_1, s_2} \wedge g_1 \implies
 \operatorname*{\bigsymb{\bigvee}}_{x \in X}
{\everymath{\displaystyle}
\mpar{\begin{array}{l}
			\alpha_{2x} = \alpha_{1} \wedge \bigwedge_\subbox{j \in J} \beta_{2xj} = \beta_{1j} \nwedge\\[12pt]
			 g_{2x} \wedge\mathcal{R}\mpar{ s'_{1},s'_{2x}}\psubst{\psi_{2x} \uplus \psi_{1}}
		\end{array}} }
%\end{array} 
\end{multline*}
\item [(3)] Deadlock reduction:
\[
\forall \mpar{s_1, s_2} \!\in\! S_1 \!\times\! S_2, V_1 \uplus V_2  \!\vdash\! \mathcal{R}\mpar{s_1, s_2} \implies( \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} ) \lor¢ \lnot( \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2}  )
\]
If there is a hole equal simulation from $A_1$ to $A_2$, then we say that  $A_2$ simulates $A_1$; we  denote it $A_2\leq A_1$
\end{definition}

 %Note also the second part of the conjunction of condition (2) expresses the deadlock reduction between automata (as introduced in Definition \ref{def:dpwd}).   
The first and second conditions coincide with the natural way to prove inductively that an automaton simulates another by starting with the initial state. The third condition ensures that  simulation  prevents the introduction of deadlocks.
Our definition captures a simple simulation for OA with the same holes that is more expressive than a strict simulation since it matches a transition with a family of transitions. 
For example,
With such a relation we are able to check the simulation between two OA that differ by duplicated states, removed duplicated transitions,  reinforced  guards, different variables, etc.
We will show in Section~\ref{sec:prop} that this simulation relation has good properties in terms of transitivity, compositionality, and reflexivity.

@@

\begin{example} \label{example:fifo-system-one}
To illustrate the simulation simulation of OA, we consider a variation on the \texttt{prod-cons} example. Namely,  we suppose that  the two processes  \texttt{P} and  \texttt{Q} communicate through a one-place  buffer. Fig. \ref{Fig:SpecOA} shows the OA modelling this simpler version of the system, that we refer to as \texttt{simprod-cons}. 
\begin{figure}[!b]
 \centering
   \includegraphics[width=.90\textwidth]{Figures/FIFOpen2.pdf}
   \caption{  The \texttt{simprod-cons} OA: the system using one-place  buffer.  \label{Fig:SpecOA}}
\end{figure}
We can easily check that this automaton simulates the one of Fig.~\ref{Fig:RefineOA}.
Indeed, one can see that  $\mathcal{R}=\{(r_0,s_0) \mapsto \texttt{l}=\texttt{f}, (r_0,s_1) \mapsto \texttt{f}=\texttt{l}+1\% \texttt{N} \}$ is  a candidate simulation 
relation. It follows that  \texttt{prod-cons} $\leq$ \texttt{simprod-cons}. \qed
\end{example}

@@

The simulation relation defined above is insufficient in the setting of composition which is the main advantage of the OA-based approach. Indeed, it should be possible to refine an automaton by filling its hole, providing a concrete view of a part of the application that was not specified originally. 
More generally, it should be possible to relate automata that do not have the same holes because composition is a crucial part of system specification.
%Thus, we believe composition should also be a form of refinement and call this feature \emph{refinement through composition}.
However, filling holes can result in a system with more or less holes than the original system because the plugged subsystem can contain itself many holes.
Next section  defines  a more powerful simulation relation able to reason on automata with different sets of holes.

@@

\section{A Simulation Relation that Takes Holes into Account}\label{sec:holes}

This section  extends the preceding relation to automata where the set of holes is not the same. Indeed, when composing the set of  holes changes, and for example we want to state whether the automaton after composition is a simulation of the original automaton or not.

@@

Unfortunately, the deadlock reduction property in the definition of simulation is not compositional: the composition operator can itself introduce a deadlock. In other words, when filling the hole of two related automata with a third one, even if there is a deadlock reduction between the two original automata, there might not be a deadlock reduction in the composed ones. The same problem may arise when two related automata are composed in the same hole of a third one.

@@

%\begin{definition}[Non-blocking composition] \label{Def:Non-block}
%Let \(A_i = \OAg[i]\) with \( 0 \leq i \leq n\) be a family of open automata.
%Let \(A = A_0\mdbrk{j_i \mapsto A_i \middle| 1 \leq i \leq n}\) and $A= \OAg$.
%
%The composition \(A_0\mdbrk{j_i \mapsto A_i \middle| 1 \leq i \leq n}\) is non-blocking if   \(A \) has a reachability predicate such that, for each reachable configuration, if there is a possible transition in \(A_0\) then there is a possible transition in \(A\):
%\begin{multline*}
% \forall s = \mpar{s_0, s_1,..,s_n} \in S, V \uplus \biguplus_\subbox{t_0 \in \fOT{s_0}} \fvars{t_0} \vdash \\ \left(\reach{A}\mpar{s} \wedge \bigvee_\subbox{t_0 \in \fOT{s_0}} \fguard{t_0} \implies \bigvee_\subbox{t \in \fOT{s}} \fguard{t} \right)
%\end{multline*}
%\end{definition}
\begin{definition}[Non-blocking composition] \label{Def:Non-block}
Let \(A_1 \defobject \OAg[1]\) and \( A_2 \defobject \OAg[2]\) be two OA.
Let $A= \OAg$ be the OA resulting from the composition \(A = A_1\subst{A_2}{k}\). 
The composition \(A_1\subst{A_2}{k}\) is non-blocking if   \(A \) has a reachability predicate such that, for each reachable configuration, if there is a possible transition in \(A_1\) then there is a possible transition in \(A\):
\begin{multline*}
 \forall s = \mpar{s_1, s_2} \in S, V \uplus \biguplus_\subbox{t \in \fOT{s_1}} \fvars{t} \vdash  \left(\reach{A}\mpar{s} \wedge \bigvee_\subbox{t \in \fOT{s_1}} \fguard{t} \implies \bigvee_\subbox{t \in \fOT{s}} \fguard{t} \right)
\end{multline*}
\end{definition}
Like in the definition of simulation (Definition~\ref{def:HoleEqualSim}) we use guards to ensure that the transition can occur. 
In general, one would not want to only consider non-blocking composition as it may reveal a bit restrictive, but it is the best necessary condition that can be identified for compositionality of simulation. It will be used to prove composition theorems in  Section~\ref{sec:prop}. 
Proving simulation relation in absence of non-blocking composition may also checked specifically for the composed automaton under study.
%It is not sufficient to ensure the existence of equivalent transitions in general, but it will be sufficient in the context of refinement.

\subsection{Automaton simulation}

One major challenge in the extension of simulation to different sets of holes is to maintain a form of transitivity while being able to take into account the actions of some of the holes. A naive definition of simulation would ensure that only the holes that are identical in the two OA are taken into account in the simulation. Unfortunately considering all the common holes does not ensure transitivity of the simulation for the following reason. If $A_1$ simulates $A_2$ and $A_2$ simulates $A_3$, and one hole $j$ appears in $A_3$ and in $A_1$ but not in $A_2$ then we have no guarantee on the way $A_1$ and $A_3$ take the actions of this his hole into account, thus  a simulation between and $A_1$ and $A_3$ would require conditions involving actions of the hole $j$ which cannot be ensured. The way we solve this issue is to remember in the simulation relation which holes have been compared. This makes the relation parameterized by a subset of the set of holes that belong to the two automata that we want to take into account.
This way, in the example above, we would have no guarantee on actions the hole $j$ by transitivity but can state a simulation relation with guarantees on the actions of the other holes.
